---
title: "Attrition_Analysis"
author: "Neddy Nyatome"
date: "4/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
DDSAnalytics is an analytics company that specializes in talent management solutions for Fortune 100 companies. Talent management is defined as the iterative process of developing and retaining employees. It may include workforce planning, employee training programs, identifying high-potential employees and reducing/preventing voluntary employee turnover (attrition). To gain a competitive edge over its competition, DDSAnalytics is planning to leverage data science for talent management. The executive leadership has identified predicting employee turnover as its first application of data science for talent management. Before the business greenlights the project, they have tasked our data science team to conduct an analysis of existing employee data.

A dataset was privided with the request to do an analysis to identify factors that lead to attrition.  Our team has focused on identifying 3 top three factors that contribute to turnover.

Additionally, we have identified some interest job role trends specific to to the roles as the business was also interested in learning about any job role trends that may exist in the data set
## Packages 
```{r}
library(regclass)
library(class)
library(caret)
library(e1071)
library(tidyverse)
library(funModeling)
library(Hmisc)
library(dplyr)

```

## Data Import and Tidying 

You can also embed plots, for example:

```{r }
employees <- read.csv('CaseStudy2-data.csv', header = T)
head(employees)
summary(employees)
str(employees)
#check if any column does have NAs
colSums(is.na(employees))
# convert some numeric variables to factors
employees <- employees %>%
  mutate(
    JobLevel = factor(JobLevel),
    StockOptionLevel = factor(StockOptionLevel),
    TrainingTimesLastYear = factor(TrainingTimesLastYear)
  )
employee

```

##EDA 
```{r}
#basic EDA on all Employees
#Compared outputs of all 3 and Identified factors that may lead to attrition to be 'YearsAtCompany', 'JobLevel' and 'Stock Option'
basic_eda <- function(employees)
{
  glimpse(employees)
  df_status(employees)
  freq(employees)
  profiling_num(employees)
  plot_num(employees)
  describe(employees)
}
  
  basic_eda(employees)
#Stats on Employees with Attrition = Yes 
employees_attrition <- employees %>% filter(Attrition == 'Yes')
# initial checkon each variable  
plot_num(employees_attrition)
basic_eda_attrition <- function(employees)
{
 glimpse(employees_attrition)
  df_status(employees_attrition)
  freq(employees_attrition)
  profiling_num(employees_attrition)
  plot_num(employees_attrition)
  describe(employees-attrition)
}

basic_eda_attrition(employees)
#Stats on Employees with Attrition = No  
employees_stayed <- employees %>% filter(Attrition == 'No')
plot_num(employees_stayed)
basic_eda_stayed <- function(employees)
{
  glimpse(employees_stayed)
  df_status(employees_stayed)
  freq(employees_stayed)
  profiling_num(employees_stayed)
  plot_num(employees_stayed)
  describe(employees_stayed)
}

basic_eda_stayed(employees)
 
#Looking more into income 
#Monthly Income Vs Years at company based on JobLevel 
ggplot(employees, aes(MonthlyIncome, YearsAtCompany)) + geom_point(aes(color = JobRole)) + 
  scale_x_continuous("MonthlyIncome", breaks = seq(0,20000,by = 2000))+
  scale_y_continuous("YearsAtCompany", breaks = seq(0,270,by = 5))+
  theme_bw() + labs(title="Scatterplot-Income vs Years at company")

#Monthly Income Vs Years at company based on Attrition 
ggplot(employees, aes(MonthlyIncome, YearsAtCompany)) + geom_point(aes(color = Attrition)) + 
  scale_x_continuous("MonthlyIncome", breaks = seq(0,20000,by = 2000))+
  scale_y_continuous("Years A tCompany", breaks = seq(0,270,by = 5))+
  theme_bw() + labs(title="Attrition- Income vs Years at company")

#Monthly Income Vs Years Since last promotion based on JobLevel 
ggplot(employees, aes(MonthlyIncome, YearsSinceLastPromotion)) + geom_point(aes(color = Attrition)) + 
  scale_x_continuous("MonthlyIncome", breaks = seq(0,20000,by = 2000))+
  scale_y_continuous("Years Since LastPromotion", breaks = seq(0,270,by = 5))+
  theme_bw() + labs(title="Attrition-Income vs Years Since Last Promotion")

ggplot(employees, aes(MonthlyIncome, YearsSinceLastPromotion)) + geom_point(aes(color = StockOptionLevel)) + 
  scale_x_continuous("MonthlyIncome", breaks = seq(0,20000,by = 2000))+
  scale_y_continuous("Years Since LastPromotion", breaks = seq(0,270,by = 5))+
  theme_bw() + labs(title="StockLevelOption-Income vs Years at company")

#Experimentation on various vizualizations 
employees %>% ggplot(aes(x = Age, y = YearsAtCompany,color = Attrition)) + geom_point()
employees %>% ggplot(aes(x = YearsSinceLastPromotion, y = YearsAtCompany,color = Attrition)) + geom_point()
employees %>% ggplot(aes(x =YearsInCurrentRole, y = YearsAtCompany,color = Attrition)) + geom_point()

ggplot(attrition, aes(x=StockOptionLevel, fill = JobRole)) +
  theme_bw()+
  geom_histogram(binwidth = 5)
employees %>% plot(x=Department, y= JobSatisfaction)
employees %>% ggplot(aes(x = MonthlyIncome,y=JobRole, color = JobLevel)) + geom_point()
employees %>% ggplot(aes(x = MonthlyIncome,y=PercentSalaryHike, color = JobLevel)) + geom_point()
employees %>% ggplot(aes(x = MonthlyIncome,y=Attrition, color = YearsSinceLastPromotion)) + geom_point()
employees %>% ggplot(aes(x = JobLevel, fill = factor(JobLevel))) + geom_bar(stat = "count") 
employees %>% ggplot(aes(x = Gender, fill = factor(JobRole))) + geom_bar(stat = "count")
employees %>% ggplot(aes(x = Gender, fill = factor(YearsSinceLastPromotion))) + geom_bar(stat = "count")
#filtering and selecting 

```

## KNN Model - Using this model to predict attrition 
```{r}
#Normalize data
num.vars <- sapply(employees, is.numeric)
employees[num.vars] <- lapply(employees[num.vars], scale)


#create training and test sets
trainInd = sample(seq(1,30,1), .7*30)
train = employees[trainInd,]
test = employees[-trainInd,]

#predict 
set.seed(800) 
test <- 1:500
train.employees <- employees.subset[-test,]
head(train.employees)
test.employees <- employees.subset[test,]
dim(train.employees)


train.def <- employees$Attrition[-test]
head(train.def)
test.def <- as.factor(employees$Attrition[test])
dim(train.def)

#test performance 

knn.1 <-  knn(train.employees, test.employees, train.def, k=2)
table(knn.1 ,test.def)
confusionMatrix(table(knn.1 ,test.def))

```



##Regression -> Using this model to predict Mothly income 
```{r}
#create the regression Model, Adding and removing predictors based on their significance 
M <- lm(MonthlyIncome ~StockOptionLevel +
          YearsInCurrentRole +
          YearsSinceLastPromotion +
          Gender+
          JobRole +
          TotalWorkingYears +
          YearsWithCurrManager +
          MonthlyRate +
          JobLevel,
          data=employees)
summary(M)
#Calculate the MSE 
RSS <- c(crossprod(M$residuals))
MSE <- RSS / length(M$residuals)
RMSE <- sqrt(MSE)
RMSE
#Using the test file without Monthly Income 
new_employees <- read.csv('test_employees.csv', header = T)
new_employees <- new_employees %>%
  mutate(
    JobLevel = factor(JobLevel),
    StockOptionLevel = factor(StockOptionLevel),
    TrainingTimesLastYear = factor(TrainingTimesLastYear)
  )
new_employees
#predict
pred_M <- predict(M, newdata = new_employees)
head(pred_M)
#Calculate RMSE for the predicted 
RSS <- c(crossprod(M$residuals))
MSE <- RSS / length(M$residuals)
RMSE <- sqrt(MSE)
RMSE

#Write to CSV 

Monthly_income_output <- write.csv(pred_M, file = "Monthly Income prediction.csv")
Monthly_income_output


```





```{r}

```